Reanalysis Step 2b: Multiple Imputation Outcome Diagnostics

### 2025-08-29 reanalysis-2b-imputation_diagnostics.qmd
enhance

    # Clinical characteristics
    mean_CCI = mean(CCI, na.rm = TRUE),
    sd_CCI = sd(CCI, na.rm = TRUE),
    mean_DCSI = mean(DCSI, na.rm = TRUE),
    sd_DCSI = sd(DCSI, na.rm = TRUE),

@@

```{r baseline-predictors}
# Merge baseline characteristics with change scores using row_id (already added earlier)
analysis_df <- changes_df |>
  left_join(df_original |> select(row_id, CCI, DCSI, age), 
            by = "row_id")

# Compare baseline predictors by imputation group
predictor_summary <- analysis_df |>
  group_by(imputation_group) |>
  summarise(
    n = n(),
    # Correlation between CCI and change
    cor_CCI_change = cor(CCI, change_baseline_to_outcome, use = "complete.obs"),
    # Correlation between baseline VA and change  
    cor_baseline_change = cor(baseline_VA_logMAR, change_baseline_to_outcome, use = "complete.obs"),
    # Mean change by CCI tercile
    mean_change_low_CCI = mean(change_baseline_to_outcome[CCI <= quantile(CCI, 0.33, na.rm = TRUE)], na.rm = TRUE),
    mean_change_high_CCI = mean(change_baseline_to_outcome[CCI >= quantile(CCI, 0.67, na.rm = TRUE)], na.rm = TRUE),
    .groups = "drop"
  )

kable(predictor_summary |> mutate(across(where(is.numeric), ~round(., 3))),
      caption = "Relationship between baseline characteristics and vision change")

# Regression model to check if group differences persist after adjustment
adj_model <- lm(change_baseline_to_outcome ~ imputation_group + CCI + DCSI + age + baseline_VA_logMAR,
                data = analysis_df)

@@

print(p_cci)
```

>>>>

    # Clinical characteristics - check if variables exist
    mean_CCI = if("CCI" %in% names(df_original)) mean(CCI, na.rm = TRUE) else NA_real_,
    sd_CCI = if("CCI" %in% names(df_original)) sd(CCI, na.rm = TRUE) else NA_real_,
    mean_DCSI = if("DCSI" %in% names(df_original)) mean(DCSI, na.rm = TRUE) else NA_real_,
    sd_DCSI = if("DCSI" %in% names(df_original)) sd(DCSI, na.rm = TRUE) else NA_real_,

@@

```{r baseline-predictors}
# Check what variables are available in the original data
available_vars <- names(df_original)
baseline_vars <- intersect(c("row_id", "CCI", "DCSI", "age"), available_vars)

# Merge baseline characteristics with change scores using row_id
if (length(baseline_vars) > 1) {
  analysis_df <- changes_df |>
    left_join(df_original |> select(all_of(baseline_vars)), 
              by = "row_id")
} else {
  # If baseline variables aren't available, create a simpler analysis
  analysis_df <- changes_df
  cat("Note: CCI, DCSI, or age not found in original data. Using simplified analysis.\n")
}

# Only do correlation analysis if CCI exists
if ("CCI" %in% names(analysis_df)) {
  # Compare baseline predictors by imputation group
  predictor_summary <- analysis_df |>
    group_by(imputation_group) |>
    summarise(
      n = n(),
      # Correlation between CCI and change
      cor_CCI_change = cor(CCI, change_baseline_to_outcome, use = "complete.obs"),
      # Correlation between baseline VA and change  
      cor_baseline_change = cor(baseline_VA_logMAR, change_baseline_to_outcome, use = "complete.obs"),
      # Mean change by CCI tercile
      mean_change_low_CCI = mean(change_baseline_to_outcome[CCI <= quantile(CCI, 0.33, na.rm = TRUE)], na.rm = TRUE),
      mean_change_high_CCI = mean(change_baseline_to_outcome[CCI >= quantile(CCI, 0.67, na.rm = TRUE)], na.rm = TRUE),
      .groups = "drop"
    )

  kable(predictor_summary |> mutate(across(where(is.numeric), ~round(., 3))),
        caption = "Relationship between baseline characteristics and vision change")
} else {
  # Simplified summary without CCI
  predictor_summary <- analysis_df |>
    group_by(imputation_group) |>
    summarise(
      n = n(),
      mean_change = mean(change_baseline_to_outcome, na.rm = TRUE),
      sd_change = sd(change_baseline_to_outcome, na.rm = TRUE),
      cor_baseline_change = cor(baseline_VA_logMAR, change_baseline_to_outcome, use = "complete.obs"),
      .groups = "drop"
    )

  kable(predictor_summary |> mutate(across(where(is.numeric), ~round(., 3))),
        caption = "Vision change by imputation group (CCI not available)")
}

# Regression model to check if group differences persist after adjustment
if (all(c("CCI", "DCSI", "age") %in% names(analysis_df))) {
  adj_model <- lm(change_baseline_to_outcome ~ imputation_group + CCI + DCSI + age + baseline_VA_logMAR,
                  data = analysis_df)

@@

  print(p_cci)
} else {
  # Simpler model without baseline covariates
  adj_model <- lm(change_baseline_to_outcome ~ imputation_group + baseline_VA_logMAR,
                  data = analysis_df)

  kable(adj_coef |> mutate(across(where(is.numeric), ~round(., 4))),
        caption = "Group differences after adjusting for baseline VA only")
}
```
